\section{Frequently Asked Questions}
\label{sec:faq}
Frequently used functionality, errors, and questions.
\paragraph{Question}
How do I install the model generation tools?
\paragraph{Answer}
All the modeling
and visualization tools are written in Python. See the README in the Eiger
root directory for more information on installing. 

\paragraph{Question}
How do I generate a model?
\paragraph{Answer}
All model generation, visualization, and serialization tasks are
performed through the \texttt{Eiger.py} script. Model generation tasks require
the DataCollection ID of the data within the Eiger database used to train
the model. As well there are many flags that can be set to specify
settings and procedures for building a model, experimenting on that model,
performing analyses, and visualizing results. 

This command will generate a model and do nothing else:

	\begin{quote}
	\texttt{Eiger.py -t <Training DataCollection ID>}
	\end{quote}

You can then specify a DataCollection to test the model against:

	\begin{quote}
	\texttt{Eiger.py -t <Training DataCollection ID> -e <Experiment DataCollection ID>}
	\end{quote}

A full listing of flags and their functions can be found by running the
following command:

	\begin{quote}
	\texttt{Eiger.py -h}
	\end{quote}

\paragraph{Question}
How do I visualize the performance or structure of my models?
\paragraph{Answer}
There are many analysis and visualization tools built into the
\texttt{Eiger.py} script.

To draw the scree graph illustrating the distribution of explained variance
across application principal components:

	\begin{quote}
	\texttt{Eiger.py <other flags> --plot-application-scree}
	\end{quote}

To draw a scatter graph of the predicted versus actual values for a given
experiment:

	\begin{quote}
	\texttt{Eiger.py <other flags> --plot-performance-scatter}
	\end{quote}

A full list of analysis flags and their functions can be found by running:

	\begin{quote}
	\texttt{Eiger.py -h}
	\end{quote}

\paragraph{Question}
How do I export my model for use in SST/macro?
\paragraph{Answer}
The following flag can be used to output a generated model for
consumption by SST/macro:

	\begin{quote}
	\texttt{Eiger.py <other flags> --output MODEL\_FILENAME}
	\end{quote}

